Browsernative Microservices

### 2017-01-26 build/paper.tex
pdf/tex build

Introduction: from simplicity to Microservices to extensible web
manifesto to w3c specifications \ldots{} and beyond

Der Erfinder der Programmiersprache Clojure, Rich Hickey, ist ohne
Zweifel eine Koryphäe auf seinem Gebiet, der Strukturierung von
komplexen Systemen. In einer vielbeachteten Keynote aus dem Jahr 2012
geht er auf etymologisch-philosophische Spurensuche nach dem Wort
\textbf{simplicity } aus Sicht eines Softwareentwicklers.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=48s}{Rails
  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} Das Adjektiv
\emph{simple} hat demnach seinen Ursprung im lateinischen Wort
\emph{simplex}, was soviel wie \emph{einfach} oder \emph{einzeln}
bedeutet. In Gegensatz dazu stehen Eigenschaften wie \emph{complex} oder
\emph{multiplex}. Qualitative Software ist, so Hickey, vor allem simpel
- im Prozess, im Design, in der Struktur und in der Entwicklung.

Bezeichnend für diese Keynote ist, dass sie von Rich Hickey im Rahmen
einer Webentwicklerkonferenz gehalten wurde. Software für den Browser
war (und ist) seit langer Zeit maßgeblich geprägt von
\textbf{Komplexität} auf mehreren Ebenen. Im Laufe dieser Bachelorarbeit
werden diese Stru

\begin{itemize}
\tightlist
\item
  Das \emph{Document Object Model} als Rückgrad jeder Webapplikation ist
  hierarchisch strukturiert und deren Elemente damit keinesfalls
  unabhängig in ihrer Darstellung und Reihenfolge.
\item
  JavaScript als single-threaded Skriptsprache lässt sich schlecht ihn
  ihrem Verhalten isolieren, ist Fehleranfällig und hatte lange Zeit nur
  sehr wenig idiomatische Lösungsansätze für komplexe Probleme, wie
  beispielsweise Asynchronität
\item
  CSS ist geprägt von stetigem Überschreiben vorher definierter Regeln
  und verletzt damit Simplizität in ihrer Struktur, Design und dem
  Entwicklungsprozesses auf bester Art und Weise.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

Betrachtet man andere populäre Webframeworks dieser Zeit, wie
beispielsweise React oder Angular, lässt sich ziemlich schnell ein
gemeinsames Designpattern ausmachen, wie Komponenten intern ihre
Zuständen verwalten. Der Facebook Entwickler Dan Abramov hat dieses
dichotome Pattern systematisch erfasst und in zwei Kategorien
eingeteilt.

Nach Abramov existieren zum einen Komponenten, die alleine für das
\textbf{Darstellen von Information} zuständig sind. Diese Komponenten
nehmen keinerlei Einfluss auf Informationen oder anderen Komponenten um
sich herum. Sie sind im wahrsten Sinne passiv und fremdgesteuert über
klar definierte Schnittstellen. Diese Komponenten sind häufig flexibel
einsetzbar und hochgradig wiederwerwendbar. Abramov nennt diese
Komponenten ``Presentational Components''.{[}1{]} Verortet man diese Art
von Komponenten innerhalb des \emph{MVC Pattern}, sind die Komponenten
reine \emph{View-Elemente}.

Im Gegensatz dazu stehen Komponenten, die für die \textbf{Verarbeitung
von Informationen} zuständig sind. Diese so genannten ``Container
Components'' haben oft einen internen Zustand, den sie verändern können
und an ihre Kindkomponenten weiterreichen können.{[}1{]} Im \emph{MVC
Pattern} handelt es sich um die \emph{Controller}. In der funktionalen
Programmiersprache Elm werden diese Elemente \emph{Updater} genannt, was
ihre Funktion noch besser umschreibt.

@@

Opening up the case for \emph{Browsernative Microservices} brings up the
question about the concept microservices in general. In fact the concept
of microservices has many facets, stretching beyond disciplines and
technical boundaries. It lacks a formal standardization but there are
certain ideas emergine from this pattern. As a primary source of truth
this articles relies on the work of Sam Newman, who has written a
comprehensive guide in \emph{Building Microservices}. The purpose of
this section is to match those ideas against the manifestations of web
components.

In a nutshell a microservice is a small, autonomous service that works
together with other services seamlessly.{[}2, p. 2{]} or with the words
of Fowler and Lewis: ``It is an approach to developing a single
application as a suite of small services, each running in its own
process and communicating with lightweight mechanisms,\ldots{}''{[}3{]}
Microservices incorporate many ideas, like \emph{domain-driven design}
where we try to represent the real world in our code.{[}2, p. 2{]} Or
making use of \emph{continuous delivery} for pushing software rapidly
through \emph{automated deployment} mechanisms in production.{[}3{]}
And, last but not least, microservices utilizes the idea of small teams
with a lot of product knowledge working mostly autonomous on their very
own service with their very own set of tools and techniques.

\subsection{Technical perspective}\label{technical-perspective}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Querying the template node in question}\\
  const node = document.querySelector(`template');
\item
  \textbf{Parsing the content and preparing the templates' content}\\
  const content = node.content;\\
  -\textgreater{} Returns a \emph{DocumentFragment} object. Handling is
  straight forward\\
  content.querySelector(`img').src = `logo.png';
\item
  \textbf{Optional: Cloning the \emph{DocumentFragment} for multiple
  use}\\
  const clone = content.cloneNode(``deep'');
\item
  \textbf{Appending the clone/original to destination}\\
  document.body.appendChild(clone);
\end{enumerate}

@@

\KeywordTok{<script>}
  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,}
    \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
     \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
       \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
       \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
       \KeywordTok{const} \NormalTok{template }\OperatorTok{=}
         \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
       \KeywordTok{this}\NormalTok{.}\AttributeTok{shadowRoot}
         \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{template}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
     \OperatorTok{\}}
   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

@@

This again is the imperative way to handle a generic \emph{HTML Import}.
In the declarative world of \emph{web components} a component is
activated, parsed and anchored solely by its' tag name
\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
Preliminary, the component needs proper configuration as the last
\texttt{HelloWorld} example wouldn't work like it is currently. The next
section will provide a better understanding about the right
configuration and composition of a component to work out-of-the-box.

@@

Despite the discrepancies among browser vendors \emph{HTML Imports}
should still be part of the paper and future \emph{web components} as no
other browser technology can bundle up CSS, JS and HTML that efficient.
As of today, January 2017, only Googles Chrome and related Opera browser
supporting the full spec and despite from \emph{HTML Imports} all other
browser vendors most likely will catch up within this year. In the
meantime, the full \emph{web components} standard can be
\textbf{polyfilled} and used across all browsers.

\subsection{Appendix: Custom Events}\label{appendix-custom-events}

@@

@@

\hypertarget{refs}{}
\hypertarget{ref-Abramov2015}{}
{[}1{]} D. Abramov, ``Presentational and Container Components --
Medium.'' 2015 {[}Online{]}. Available:
\url{https://medium.com/@dan_abramov/smart-and-dumb-components-7ca2f9a7c7d0}.
{[}Accessed: 01-Dec-2016{]}

\hypertarget{ref-Newman2015}{}
{[}2{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
2016 {[}Online{]}. Available:
\url{http://www.ebook.de/de/product/22539693/sam_newmann_building_microservices.html}

\hypertarget{ref-Fowler2014}{}
{[}3{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
new architectural term,'' Jan. 2014 {[}Online{]}. Available:
\url{http://www.martinfowler.com/articles/microservices.html}

\hypertarget{ref-Martin}{}
{[}4{]} R. C. Martin, ``The single responsibility principle.''
{[}Online{]}. Available:
\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}

\hypertarget{ref-Conway1968}{}
{[}5{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
Available: \url{http://www.melconway.com/Home/Committees_Paper.html}

\hypertarget{ref-Issa2016}{}
{[}6{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
{[}Online{]}. Available:
\url{https://www.youtube.com/watch?v=8ZTFEhPBJEE}

\hypertarget{ref-Buchner2016}{}
{[}7{]} D. Buchner, ``Demythstifying web components,'' 2016
{[}Online{]}. Available:
\url{http://www.backalleycoder.com/2016/08/26/demythstifying-web-components/}

\hypertarget{ref-vanKesteren2014}{}
{[}8{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
{[}Online{]}. Available:
\url{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}

\hypertarget{ref-Bidelman2016}{}
{[}9{]} E. Bidelman, ``Custom elements v1: reusable web components.''
2016 {[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/primers/customelements/}.
{[}Accessed: 01-Dec-2016{]}

\hypertarget{ref-HTML}{}
{[}10{]} \emph{HTML living standard --- last updated 11 january 2017}.
{[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}

\hypertarget{ref-Bidelman2016shadow}{}
{[}11{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
2016 {[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/getting-started/primers/shadowdom}

>>>>

Computers can scale, humans can't. Ever since a program or complex
system made by humans has been constrained by humans mental
capabilities. Like in the analogy of juggling balls, our brain can just
``juggle'' a few things at a time. Rich Hickey, the inventor of the
programming language \emph{Clojure} gave an inspirational keynote on the
topic of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
of a humans life, simplicity aligns perception with our mental
capacities. Understanding the word \textbf{simple} alone will outline
the this paper on its own.

Derived from the ancient Latin word \textbf{simplex}, simple can be
understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
  Dictionary}} which points directly to the unidimensional aspect. While
complexity describes the multilayered und entangled nature of
conditions, simplicity empowers the human brain to reason about issues
in a straightforward manner. It certainly has some overlapping's with
\emph{easy}, but while \emph{easy} is more of a relative nature, simple
can be laid out as a objective manner - a goal software development
should pursuit and praise.

Software development is undoubtedly rich in complexity and full of
subtle pitfalls for the human brain. In a typical scenario, a piece of
software evolves over time in one or another \sout{opinionated}
direction. Layers of new abstractions wrestling with old legacy
abstractions and mutating objects will get out of control. Subtle bugs
start to creep in. Eventually the small piece of software ends up in a
highly complected monolith which will determine future design decisions
to a painful degree. Any company will be highly determined by the
current state trying to ``keep the lights on''.

On the other hand, any truly agile system architecture is laid out in a
fine-grained manner. As Rich Hickey argues, design decisions should be
made under the \textbf{impression of extending, substitution, moving,
combining and repurposing}. The ability to reason about the program at
any given time is crucial for future decisions and implementations.
Recalling again the unidimensional nature of simplicity.

The \emph{state of the web} is certainly a different kind of complex
multidimensional beast. ``The web'' is coined to be meant ``everything
that runs in the browser''. While simplicity in the backend is mostly a
matter of principles, any frontend developer is highly restricted on the
entangled nature of the browser platform.

In the last four years the average transfer size of a webpage doubled to
currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
  Trends}} Subtracting images, fonts or other content the size of HTML,
CSS and JS sums up to a total average of 550 kb. One character weights
around 1 byte which means an average webpage is delivering 550.000
character or around 125 pages of single-spaced text. Frederic Filloux
calculated the ratio of real content on different newspaper websites and
came to the conclusion, that only round about 5-6 \% of the transferred
characters made for consumption.{[}1{]}

Having an 95 \% overhead is rather undesirable for both the consumers
and creators of the website. Since it's a widespread problem without a
single point of failure one can argue the platform itself is the
failure. By design, every pageload results in a \textbf{global DOM tree}
managed by the browser engine. Regardless of rendering just a bunch of
static text nodes or an ever changing webapp the underlying global model
remains the same. \textbf{Encapsulation and modularization} is a typical
pattern in the computer industry to make complexity manageable and to
accommodate future uncertainty.{[}2, p. 1{]} Today many websites tend
towards becoming complex single-page applications (and therefore the
average JS payload is soaring too). Consequently, the demands to the
browser platform changed from a static page renderer to a
\textbf{complexe UI machine}. Without additional layers of abstraction
modularization would be impossible to archive.

In the recent years many \textbf{frameworks}, libraries and
methodologies tried to tackle the global nature of the DOM by scoping
assets and design rules into maintainable components. While the DOM
can't be scoped, JS can. Therefore many of those frameworks, like
ReactJS, AngularJS or VueJS just to name a few, ditched the old rule of
separated HTML, CSS and JS in favor of an additional layer of
\textbf{abstracted JS components} (containing content, markup and
styling). Quiet often those frameworks mimic a MVC pattern on top of the
browser engine which is a reasonable simple design pattern to build
graphical user interfaces. This paper will apply the same pattern for
building a browsernative microservice.While frameworks are a valid
approach to build scalable web applications they remain highly
opinionated and can change and break over time. Furthermore, they
inflate the code with libraries as well. A standardized way for creating
complex UI interfaces painlessly requires new build-in browser
capabilities.

In 2013 thinkers, creators and browser vendors joined together to
propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
  Extensible Web Manifesto}} The claim of the Manifesto is to enhance
the current web platforms with new low-level capabilities. Those
features should empower creators of the web to write more declarative
code and therefore overcome known bottlenecks and artificial
abstractions. Four years later, the enhancement of JavaScript
leapfrogged and many new low-level APIs brought to life. With the new
APIs at hand a webdeveloper can create scoped and highly reusable
microservices without additional libraries even directly in the browser
console.

\emph{Disclaimer:} Currently, all new technologies proposed throughout
this paper can be tried and tested in Google Chrome, Opera and the
Safari Technology Preview. Usually the browser adoption time is short in
time an therefore full support of all standards is expected soon.
Meanwhile this standards can be used through \textbf{polyfills} on every
important legacy browser.

@@

Calling the need for \emph{Browsernative Microservices} brings up the
question about the concept of microservices in general. In fact the
concept of microservices has many facets, stretching beyond disciplines
and technical boundaries. It lacks a formal standardization but subsumes
certain ideas emergine from this pattern. As a primary source of truth
this paper relies on the work of Sam Newman, who has written a
comprehensive guide in \emph{Building Microservices} and the work of
Fowler and Lewis. The purpose of this section is to match those ideas
against the browser platform where different rules of development
applies.

In a nutshell a microservice is a small, autonomous service that works
together with other services seamlessly.{[}3, p. 2{]} or with the words
of Fowler and Lewis: ``In short, the microservice architectural style is
an approach to developing a single application as a suite of small
services, each running in its own process and communicating with
lightweight mechanisms, often an HTTP resource API''{[}4{]}
Microservices incorporate many other ideas, like \emph{domain-driven
design} where we try to represent the real world in our code.{[}3, p.
2{]} Or making use of \emph{continuous delivery} for pushing software
rapidly through \emph{automated deployment} mechanisms in
production.{[}4{]} And, last but not least, microservices utilizes the
idea of small teams with a lot of product knowledge working mostly
autonomous on their very own service with their very own set of tools
and techniques.

\subsection{Componentization via
Services}\label{componentization-via-services}

``A~\textbf{component}~is a unit of software that is independently
replaceable and upgradeable.''{[}4{]} Compared to libraries a component
service offers multiple advantages. A library is only loosely coupled to
the implementation and therefore hard to track in functionality.
Changing a library often requires an unforeseen amount of time fixing
implementations. Therefore in the field we often see multiple versions
of libraries deployed simultaneous (like with JQuery) which adds a huge
amount of overhead on the page load. Eventually, dead code elimination
won't work in highly coupled libraries. Components instead incorporate a
much better evolution mechanism in the service contracts. Changing
functionality won't break other components. A component can
progressively enhanced which guarantees functionality throughout
different versions.

Compared to libraries componentized services offers a more explicit
interface.{[}4{]} While the functionality of a library needs
documentation, a component functionality is exposed via the components'
signature and therefore highly accessible. HTML elements are usually
controlled via their attributes. A generic HTML service therefore might
look like this:
\texttt{\textless{}my-button\ isPrimary\textgreater{}\textless{}/button\textgreater{}}.

\subsection{Organized around Business
Capabilities}\label{organized-around-business-capabilities}

\subsection{Products not Projects}\label{products-not-projects}

\subsection{Smart endpoints and dumb
pipes}\label{smart-endpoints-and-dumb-pipes}

\subsection{Decentralized Governance}\label{decentralized-governance}

\subsection{Decentralized Data
Management}\label{decentralized-data-management}

\subsection{Infrastructure Automation}\label{infrastructure-automation}

\subsection{Design for failure}\label{design-for-failure}

\subsection{Evolutionary Design}\label{evolutionary-design}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Querying the template node in question}\\
  const node = document.querySelector(`template');
\item
  \textbf{Parsing the content and preparing the templates' content}\\
  const content = node.content;\\
  -\textgreater{} Returns a \emph{DocumentFragment} object. Handling is
  straight forward content.querySelector(`img').src = `logo.png';
\item
  \textbf{Optional: Cloning the \emph{DocumentFragment} for multiple
  use}\\
  const clone = content.cloneNode(``deep'');
\item
  \textbf{Appending the clone/original to destination}\\
  document.body.appendChild(clone);
\end{enumerate}

@@

\KeywordTok{<script>}
 \CommentTok{// Switched to anonymous class notation}
 \CommentTok{// for keeping associated code together.}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,}
  \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
   \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
    \KeywordTok{const} \NormalTok{template }\OperatorTok{=}
     \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{shadowRoot}
     \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{template}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
   \OperatorTok{\}}
  \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

@@

This again is the imperative way to handle a generic \emph{HTML Import}.
In the declarative world of \emph{web components} a component is
activated, parsed and anchored solely by its' tag name
\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
Preliminary, the component needs proper configuration as the last
\texttt{HelloWorld} example wouldn't work like it is currently. The next
section will elaborate the right configuration and composition of a
component to work out-of-the-box.

@@

Despite the discrepancies among browser vendors \emph{HTML Imports}
should still be part of the paper and future \emph{web components} as no
other native browser technology can bundle up CSS, JS and HTML that
efficient. As of today, January 2017, only Googles Chrome and related
Opera browser supporting the full \emph{web components} spec and,
despite from \emph{HTML Imports}, all other browser vendors most likely
will catch up with \emph{Custom Elements} and \emph{Shadow DOM} within
this year. In the meantime, the full \emph{web components} stack can be
\textbf{polyfilled} and used across all browsers.

\subsection{\texorpdfstring{Appendix A: Custom Events
\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{Appendix A: Custom Events (whatwg)}}\label{appendix-a-custom-events-whatwg}

Events are first-class citizens in the browser world and \emph{Custom
Events} are no exception. The \emph{Custom Elements} interface is part
of the DOM since years but with the rise of \emph{Custom Elements} they
will most likely become an indispensable building block of \emph{web
components}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> index.html}
\KeywordTok{<hello-world>}
 \KeywordTok{<button>}\NormalTok{Launch CustomEvent}\KeywordTok{</button>}
\KeywordTok{</hello-world>}
\CommentTok{<!-- COMPONENT STARTS HERE -->}
\KeywordTok{<script>}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} 
 \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
     \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
   \CommentTok{// Craft a CustomEvent e}
     \KeywordTok{const} \NormalTok{e }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \OperatorTok{\{}
    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,} \CommentTok{//important!}
    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains string or object'}
   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
   \CommentTok{// Launch e on child button click}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(e)}
    \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
   \CommentTok{// Catch e. Typically done by some parent}
     \CommentTok{// node. Message in detail property}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\VariableTok{e}\NormalTok{.}\AttributeTok{detail}\NormalTok{)}
   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
  \OperatorTok{\}}
 \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

Naming events after the emitting tag makes the API almost
self-explanatory. Fortunately the \texttt{detail} property can transmit
even objects witch allows the developer to craft almost all
functionality inside one event. The further sections will elaborate a
feasible architecture for building a scaling microservice architecture.

Chaining and aggregating events from child nodes should be practiced and
exercised quiet frequently. As mentioned earlier in the \emph{Custom
Elements} section, the pattern of \textbf{extending native elements}
should be somewhat dismissed as it may be implemented outside of Chrome.
Nevertheless, it is possible to create own kind of quasi native buttons
when chaining a \emph{CustomEvent} directly after the native click
event.

Another typical \emph{custom element} use case can be as an actor on
(native) child elements. In this case, the \emph{Custom Element} catches
events from children, buffers or rebuild them and eventuall fires an
event towards the document root.

Unfortunately events only work ``upstream'' towards parent nodes. Still
the web platform offers plenty of possibilities to talk back to child
nodes.

\subsection{\texorpdfstring{Appendix B: Web Worker
\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{Appendix B: Web Worker (whatwg)}}\label{appendix-b-web-worker-whatwg}

Like \emph{Custom Events}, \emph{Web Workers} had been around for a long
time and therefore embrace full support among major browsers. They
emerged at around 2009 when discussions about browser performance was
still in the early days. Nevertheless, the addressed problem of
\emph{Web Workers} is a fundamental language problem of JS itself.

JS runs in a single-threaded language environment. Every script in the
browser environment, from handling UI events to query and process larget
amounts of API data and manipulating the DOM, runs on the same
thread{[}13{]}. Putting a lot of work to the single main thread can slow
down the web service significantly. From time to time scripts can block
or fail to whatever reason which leads to a frozen UI on the users side.
A worker can overcome the bottleneck of the single-threaded nature with
spawning a new \textbf{background thread}. Most of the browsers work can
be leveraged to this new thread. Todays web architectures aims to
leverage an increasing amout of proccessing to the client to avoid
time-consuming roundtrips especially in mobiles networks.\footnote{Latency
  numbers: https://gist.github.com/jboner/2841832} While many native
APIs like \texttt{fetch} work seamlessly in the new thread, a worker has
no access to the DOM at all.

A \emph{Web Worker} spawns a new background thread where scripts can run
concurrent to the main thread. Usually a worker is loaded from a workers
dedicated file to embrace this kind of separation.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{const} \NormalTok{worker }\OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'worker.js'}\NormalTok{)}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

After initialization a worker communicates over a simple \textbf{message
based interface} with the main thread.

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
\CommentTok{// Send to worker}
\VariableTok{worker}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{'Hello World'}\NormalTok{)}\OperatorTok{;}
\CommentTok{// Receive msg from worker}
\VariableTok{worker}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
 \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\StringTok{'Worker said: '}\OperatorTok{,} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{worker}\NormalTok{.}\AttributeTok{js}
\CommentTok{// Receive msg and echo back}
\KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
 \KeywordTok{this}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{"Echo "} \OperatorTok{+} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

Having no access to the DOM can be seen as hinderance,but i

@@

dichotome Pattern

@@

\hypertarget{refs}{}
\hypertarget{ref-Filloux2016}{}
{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
{[}Online{]}. Available:
\url{https://mondaynote.com/bloated-html-the-best-and-the-worse-cac6eb06496d}

\hypertarget{ref-Baldwin2006}{}
{[}2{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the Design of
Complex Engineering Systems,'' in \emph{Complex engineered systems:
Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205
{[}Online{]}. Available:
\href{http://dx.doi.org/10.1007/3-540-32834-3\%7B/_\%7D9}{http://dx.doi.org/10.1007/3-540-32834-3\{\textbackslash{}\_\}9}

\hypertarget{ref-Newman2015}{}
{[}3{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
2016 {[}Online{]}. Available:
\url{http://www.ebook.de/de/product/22539693/sam_newmann_building_microservices.html}

\hypertarget{ref-Fowler2014}{}
{[}4{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
new architectural term,'' Jan. 2014 {[}Online{]}. Available:
\url{http://www.martinfowler.com/articles/microservices.html}

\hypertarget{ref-Martin}{}
{[}5{]} R. C. Martin, ``The single responsibility principle.''
{[}Online{]}. Available:
\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}

\hypertarget{ref-Conway1968}{}
{[}6{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
Available: \url{http://www.melconway.com/Home/Committees_Paper.html}

\hypertarget{ref-Issa2016}{}
{[}7{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
{[}Online{]}. Available:
\url{https://www.youtube.com/watch?v=8ZTFEhPBJEE}

\hypertarget{ref-Buchner2016}{}
{[}8{]} D. Buchner, ``Demythstifying web components,'' 2016
{[}Online{]}. Available:
\url{http://www.backalleycoder.com/2016/08/26/demythstifying-web-components/}

\hypertarget{ref-vanKesteren2014}{}
{[}9{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
{[}Online{]}. Available:
\url{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}

\hypertarget{ref-Bidelman2016}{}
{[}10{]} E. Bidelman, ``Custom elements v1: reusable web components.''
2016 {[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/primers/customelements/}.
{[}Accessed: 01-Dec-2016{]}

\hypertarget{ref-HTML}{}
{[}11{]} \emph{HTML living standard --- last updated 11 january 2017}.
{[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}

\hypertarget{ref-Bidelman2016shadow}{}
{[}12{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
2016 {[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/getting-started/primers/shadowdom}

\hypertarget{ref-Bidelman2010}{}
{[}13{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}
